---
title: "COVID-19 in the Western Balkans"
author: "Imer Muhovic"
date: "`r Sys.Date()`"
output: html_document
---

This page contains a number of visualizations of COVID19 cases in the Western Balkans, it is hosted on GitHub pages, and all the code and graphs are free to use, share, modify any way you see fit with proper attribution. [Done with the support of Balkansmedia.](http://balkansmedia.org/)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggthemes)
library(plotly)
library(tibbletime)
library(zoo)
```



```{r overall_confirmed, include=FALSE}

time_series_19_covid_Confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))

confirmed_it_100 <- time_series_19_covid_Confirmed %>% select(Country  = `Country/Region`, everything(), -Lat, -Long, -`Province/State`) %>% pivot_longer(-Country, "Date") %>% filter(Country %in% c('Serbia', "Bosnia and Herzegovina", "Croatia", "Slovenia", "North Macedonia", "Kosovo", "Albania", "Montenegro", "Italy")) %>% filter(value > 100)
confirmed_it_100$Date <- confirmed_it_100$Date %>% as.Date(format="%m/%d/%y")

confirmed <- time_series_19_covid_Confirmed %>% select(Country  = `Country/Region`, everything(), -Lat, -Long, -`Province/State`) %>% pivot_longer(-Country, "Date") %>% filter(Country %in% c('Serbia', "Bosnia and Herzegovina", "Croatia", "Slovenia", "North Macedonia", "Kosovo", "Albania", "Montenegro", "Italy")) %>% filter(value > 0)
confirmed$Date <- confirmed$Date %>% as.Date(format="%m/%d/%y")

overall_confirmed_graph <- confirmed_it_100 %>% group_by(Country) %>% mutate(days_since_start = difftime(Date, min(Date), units = "days")) %>% filter(days_since_start < 28) %>% ggplot(aes(days_since_start, value, color = Country)) + geom_line(size = 2) + theme_wsj() + theme(legend.text = element_text(size = 12), plot.title = element_text(size = 14, hjust = 0.5), plot.subtitle = element_text(size = 10)) + scale_color_brewer(palette = "Paired") + labs(title = "Number of cases of COVID19 in Western Balkans \n since 100th reported case", caption ="Data source:https://github.com/CSSEGISandData/COVID-19/", subtitle = "By: Imer Muhovic @imerm1", x = "Days since 100th confirmed case") + scale_y_log10()
```

```{r, include=FALSE}

rolling_avg <- rollify(mean, window = 7)

rolling_daily_it_30 <- time_series_19_covid_Confirmed %>% select(Country  = `Country/Region`, everything(), -Lat, -Long, -`Province/State`) %>% pivot_longer(-Country, "Date") %>% filter(Country %in% c('Serbia', "Bosnia and Herzegovina", "Croatia", "Slovenia", "North Macedonia", "Kosovo", "Albania", "Montenegro", "Italy"), value > 0) %>% group_by(Country) %>% filter(n() > 7) %>% mutate(Date = as.Date(Date, format="%m/%d/%y"), new_cases = value - dplyr::lag(value, n = 1)) %>% mutate(avg = rolling_avg(new_cases)) %>% filter(ifelse(new_cases >9, TRUE, NA) %>% na.locf(na.rm=FALSE)) %>% filter(!is.na(avg)) %>% ungroup()

rolling_graph_static <- rolling_daily_it_30 %>% group_by(Country) %>% mutate(days_since_start = difftime(Date, min(Date), units = "days")) %>% filter(days_since_start < 28) %>% ggplot(aes(days_since_start, avg, color = Country)) + geom_line(size = 2) + theme_wsj() + theme(legend.text = element_text(size = 12), plot.title = element_text(size = 14, hjust = 0.5), plot.subtitle = element_text(size = 10), plot.caption = element_text(size = 8)) + scale_color_brewer(palette = "Paired") + labs(title = "Average of daily reported cases of COVID19 in Western Balkans", caption ="Data source:https://github.com/CSSEGISandData/COVID-19/", subtitle = "By: Imer Muhovic @imerm1, Y axis is average of previous 7 days", x = "Days since 100th confirmed case") + scale_y_log10()

rolling_graph_interactive <- rolling_daily_it_30 %>% group_by(Country) %>% mutate(days_since_start = difftime(Date, min(Date), units = "days")) %>% filter(days_since_start < 28) %>% ggplot(aes(days_since_start, avg, color = Country)) + geom_line(size = 2) + theme_clean()+ scale_color_brewer(palette = "Paired") + scale_y_log10() + labs(title = "Daily average new cases for past week") + xlab("Days since 10 cases per day were first reported") + ylab("Average cases reported over past 7 days")

```




```{r recovered, include=FALSE}
time_series_19_covid_recovered <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv"))

recovered <- time_series_19_covid_recovered %>% select(Country  = `Country/Region`, everything(), -Lat, -Long, -`Province/State`) %>% pivot_longer(-Country, "Date") %>% filter(Country %in% c('Serbia', "Bosnia and Herzegovina", "Croatia", "Slovenia", "North Macedonia", "Kosovo", "Albania", "Montenegro")) %>% filter(value > 0)
recovered$Date <- recovered$Date %>% as.Date(format="%m/%d/%y")

recovered_graph <- recovered %>% group_by(Country) %>% mutate(days_since_start = difftime(Date, min(Date), units = "days")) %>% ggplot(aes(days_since_start, value, color = Country)) + geom_line(size = 2) + theme_wsj() + theme(legend.text = element_text(size = 12), plot.title = element_text(size = 14, hjust = 0.5), plot.subtitle = element_text(size = 10)) + scale_color_brewer(palette = "Paired") + labs(title = "Number of recoveries from COVID19 in Western Balkans", caption ="Data source:https://github.com/CSSEGISandData/COVID-19/", subtitle = "By: Imer Muhovic @imerm1", x = "Days since first recovery")

```

```{r deceased, include=FALSE} 
time_series_19_covid_deaths <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))

dead <- time_series_19_covid_deaths %>% select(Country  = `Country/Region`, everything(), -Lat, -Long, -`Province/State`) %>% pivot_longer(-Country, "Date") %>% filter(Country %in% c('Serbia', "Bosnia and Herzegovina", "Croatia", "Slovenia", "North Macedonia", "Kosovo", "Albania", "Montenegro")) %>% filter(value > 0)
dead$Date <- dead$Date %>% as.Date(format="%m/%d/%y")

dead_graph <- dead %>% group_by(Country) %>% mutate(days_since_start = difftime(Date, min(Date), units = "days")) %>% ggplot(aes(days_since_start, value, color = Country)) + geom_line(size = 2) + theme_wsj() + theme(legend.text = element_text(size = 12), plot.title = element_text(size = 14, hjust = 0.5), plot.subtitle = element_text(size = 10)) + scale_color_brewer(palette = "Paired") + labs(title = "Number of deaths from COVID19 in Western Balkans", caption ="Data source:https://github.com/CSSEGISandData/COVID-19/", subtitle = "By: Imer Muhovic @imerm1", x = "Days since first death")
```

```{r totals-plot, include=FALSE}
totals_conf <- confirmed %>% filter(Country != "Italy") %>% group_by(Country) %>% top_n(1, Date) %>% ungroup() %>% select(everything(), confirmed = value, - Date)
totals_dead <- dead %>%group_by(Country) %>% top_n(1, Date) %>% ungroup() %>% select(everything(), dead = value, -Date)
totals_recov <- recovered %>%group_by(Country) %>% top_n(1, Date) %>% ungroup() %>% select(everything(), recovered = value, -Date)

total <- left_join(totals_conf, totals_dead) %>% left_join(totals_recov) %>% pivot_longer(c(confirmed, dead, recovered))
totals_graph <- ggplot(total) + geom_col(aes(x = Country, y = value, fill = name))  + scale_fill_manual(values = c("dead"="#E41A1C", "confirmed"="#377EB8","recovered"= "#4DAF4A")) + theme_clean() + labs(title = "Western Balkans COVID19 current state", fill="") + ylab("") + xlab("") + theme(axis.text.x = element_text(angle=45, hjust = 1))

```

## Daily average new cases
This graph displays the daily average in new cases for the past 7 days for each country. X axis starts with first day that 10 cases were reported in each respective country, uses a Log10 scale, with Italy added for comparison.

```{r rolling-interactive, plotly=TRUE,fig.width=9, fig.height=6, echo=FALSE}
ggplotly(rolling_graph_interactive)
```



## Overall COVID19 Growth Rate

A comparison of WB countries that have >100 confirmed cases with Italy. Please note that we are using a LOG10 scale, which is the standard for exponential growth, if the slope of the curve starts decreasing it means that the growth rate of the epidemic is decreasing as well, [as explained in this Balkansmedia video](https://www.youtube.com/watch?v=nHoJ_hICzyM&t=2s).

```{r pressure, plotly=TRUE,fig.width=9,fig.height=6, echo=FALSE}

interactive_confirmed_graph <- confirmed_it_100 %>% group_by(Country) %>% filter(value > 100) %>% mutate(days_since_start = difftime(Date, min(Date), units = "days")) %>% filter(days_since_start < 28) %>% ggplot(aes(days_since_start, value, color = Country)) + geom_line(size = 2) + scale_y_log10() + theme_clean() + scale_color_brewer(palette = "Paired") + labs(title = "Number of cases of COVID19 in Western Balkans \n since 100th confirmed case", caption ="Data source:https://github.com/CSSEGISandData/COVID-19/", x = "Days since 100th confirmed case, Italy added for comparison")
ggplotly(interactive_confirmed_graph)
```

## Overall COVID19 State In Western Balkans

```{r interactive-totals, plotly=TRUE,fig.width=9, fig.height=6}
ggplotly(totals_graph)
```


## Static Plots
If you have issues embedding the above shown code you can freely copy the images displayed below:

```{r, fig.width=12,fig.height=8}
rolling_graph_static
interactive_confirmed_graph
totals_graph
overall_confirmed_graph
recovered_graph
dead_graph
```

